### 2018-04-11 paper.Rmd
worked on technical section

<!-- Remind readers why they're reading this: understanding the structure helps you use and understand the tools -->

<!-- We first present the design and technical approach of childes-db, then describe the interactive analysis tools provided by our web-based frontend.  -->

@@

Unlike the *parsing* approach in CLAN which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts. Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges). Although we expect most users to interact with the database through the `childesr` API for the `R` language described below, we begin with a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme.

At its core, childes-db is a database consisting of a set of linked tabular data stores where rows correspond to records. <!-- what did you mean here? don't rows = records--> The smallest unit of abstraction tracked by the database is a token, treated here as the citation orthographic form of a word. Using the standardized written form of the word facilitates the computation of lexical frequency statistics, for comparison or aggregation across children or time periods. Deviations from the citation form -- reflecting the deviations from citation forms that are particularly common in the course of language development -- are kept as a separate (possibly null) field associated with each token.

@@

A *transcription* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES. A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus. Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES. Because every entity can be linked to a top-level collection, childes-db places content from all languages in the same table for a given level of abstraction <!-- a little confused by this last sentence -->.

We chose to treat participants separate from the above hierarchy because children can occur across transcripts (e.g., the Brown Corpus). A participant is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI, MOT, FAT, etc.), and the range of ages (or age of corresponding child) for which they are observed. For non-child participants, the record contains an identifier for the target child to associate caregivers and children.

@@

In order to import the CHILDES corpora into the MySQL schema described above, it must first be accurately parsed and subsequently vetted to ensure the integrity of the data migration. Along with being made available in CHAT, the corpora are also downloadable in a format known as eXtensible Markup Language, or XML; this significantly mitigated any problems related to parsing CHAT as it allowed us to leverage the built-in XML parsing libraries available in numerous programming languages such as Python. Once the XML data were crawled and parsed into tokens and utterances, they were imported into a MySQL database. The meticulously constructed CHILDES XML Schema allowed us to make precise decisions as to what information we found relevant to our database and what was to be discarded or ignored; further information on what data were retained can be found below.

The data imported into childes-db requires extensive data integrity checks to ensure that our parse of the corpora is accurate and preferable over the adhoc parsers developed by many individual researchers. In order to evaluate this, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT. <!-- technically this is all we do right now -->

<!-- [[Details of testing: Why trust our parse over your in-house library? CLAN TESTING DETAILS - also mention versioning as a reason to trust]] -->

@@

<!-- Continue to make the case that it's important to version! childes changes without warning, corpora get added, etc. -->

The content of CHILDES changes as additional corpora are added or transcriptions are updated; currently, these changes are not systematically tracked.
Here we introduce a simple versioning system by proucing a new complete parse of the current state of CHILDES every 6 months.
Users interact with the most recent version of the database available by default. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last 6 versions) through our `childesr` API so that researchers can run analyses against specific versions of the database.
<!-- Potentially put some R code to show easy it is to access previous versions? -->
For versions more than three years old, users have the options of downloading compressed .sql files and serving them with a local installation of MySQL server.
<!-- How easy is it to implement versioning in the Shiny apps?-->

@@

- Sparsely annotated tiers, e.g. phonology (`%pho`) and situation (`%sit`)
- Media links, including timing information <!-- we have media timestamps -->
- compound word markers, indicated with underscores ("green_eggs_and_ham") <!-- we have this -->
- Tone direction and stress
- Filled pauses 
- Reformulations, word revision, and phrase revision, e.g. <what did you> [//] how can you see it ?
- paralinguistic material, e.g. [=! cries]

>>>>

<!-- Remind readers why they're reading this: understanding the structure helps you use and understand the tools. SM: Moved this up from database format because it addresses more general issues -->
Unlike the *parsing* approach in CLAN which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts.
Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges).
Although we expect most users to interact with the database through the `childesr` API for the `R` language described below, we begin with a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme.
<!-- We first present the design and technical approach of childes-db, then describe the interactive analysis tools provided by our web-based frontend.  -->

@@

At its core, childes-db is a database consisting of a set of linked tabular data stores where rows correspond to linguistic entities like words, utterances, and sampling units like transcriptions and corpora. 
<!-- what did you mean here? don't rows = records; SM: that's a description of a database in general; we can make a specific assertion about childes-db-->
The smallest unit of abstraction tracked by the database is a token, treated here as the citation orthographic form of a word.
Using the standardized written form of the word facilitates the computation of lexical frequency statistics, for comparison or aggregation across children or time periods. 
Deviations from the citation form -- reflecting the deviations from citation forms that are particularly common in the course of language development -- are kept as a separate (possibly null) field associated with each token.

@@

A *transcription* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES.
A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus.
Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES.
Because every record can be linked to a top-level collection (generally corresponding to a language), each table includes data from all languages represented in CHILDES.

We chose to treat participants separate from the above hierarchy because children can occur across transcripts (e.g., the Brown Corpus).
A participant identifier is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI, MOT, FAT, etc.), and the range of ages (or age of corresponding child) for which they are observed.
For non-child participants (caregivers and others), the record additionally contains an identifier for the corresponding target child to associate caregivers and children.

@@

In order to import the CHILDES corpora into the MySQL schema described above, it must first be accurately parsed and subsequently vetted to ensure its integrity.
We parse the XML (eXtensible Markup Language) release of CHILDES hosted by \url{childes.talkbank.org} using the NLTK library in Python.
<!-- citation for NLTK -->
<!-- details of our contribution to the CHILDES corpus module in NLTK---didn't we change it? -->
Logic implemented in Python converts the linear, multi-tier parse into a tabular format appropriate for childes-db.
This logic includes decisions regarding what information sources are captured in the current release of the database, and which were left for future development (detailed below).

<!-- testing details -->
The data imported into childes-db requires extensive data integrity checks to ensure that our parse of the corpora is accurate and preferable over ad hoc parsers developed by many individual researchers.
In order to evaluate this, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT.
<!-- SM: what were the results of these comparisons?  -->
<!-- AS: technically this is all we do right now SM: what more can we do for low technical investment-->

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
Here we introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months.
By default, users interact with the most recent version of the database available. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific hsitorical versions of the database.
<!-- Potentially put some R code to show easy it is to access previous versions? -->
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions-->
<!-- SM: can people get to these sql files currently on the website?-->

@@

- Sparsely annotated tiers, e.g. phonology (`%pho`) and situation (`%sit`)
<!--- Media links, including timing information AS: we have media timestamps -->
<!--- Compound word markers, indicated with underscores ("green_eggs_and_ham") AS: we have this -->
- Tone direction and stress
- Filled pauses 
- Reformulations, word revision, and phrase revision, e.g. <what did you> [//] how can you see it ?
- paralinguistic material, e.g. [=! cries]
